Hunting experience shapes individual foraging specialisation and predator-prey interactions in an online videogame

### 2023-02-09 manuscript/Chapter2-manuscript.Rmd
update text methods and results

For each level of experience (i.e. novice, intermediate, advanced), the model estimated the mean speed of every individual (which we call the mean part of the model), and, employing a heterogeneous structure of the residuals, the intra-individual standard deviation in speed of every individual (which we call the dispersion part of the model). Thus, in the mean part of the model, we could estimate among individual differences in the foraging tactic used by predators (i.e. fast or slow), while in the dispersion part of the model, we could estimate among individual differences in foraging specialization. We followed the same structure for the preys' speed to estimate among individual differences in the mean and standard deviation in speed of the prey encountered by predators. For hunting success, we only modeled the mean part of the equation to estimate among individual differences in mean hunting success. For the predator and the prey's speed, we controlled for the average rank of the prey that they encountered on both the mean and dispersion part of the equation. The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank). We used the mean rank of the four prey and applied a square-root transformation. For the predator's hunting success, we controlled for the square-root of the duration of the match. The model has the following form, with each equation repeated three times (i.e. one for each level of experience) :

@@

At the individual level, we detected differences in the average movement speed among predators ($SD\ predator\ ID$, Table 2). These differences remained stable (although they decreased minimally) at the intermediate stage (Table 2 and Figure S4), and increased at the advanced stage to a level higher than when predators were novice (Table 2 and Figure S4). The largest proportion of variation in mean predator speed was attributed to differences among predator avatars ($ICC\ avatar\ ID$, Table S1), and these differences increased with experience ($SD\ avatar\ ID$, Table 2). The variation in mean predator speed among the virtual environments was very low ($SD\ environment\ ID$, Table 2). For the dispersion part of the model (i.e. sigma predator speed), predators displayed considerable differences in their degree of hunting specialization, which also increased with experience ($SD\ predator\ ID$, Table 2 and Figure S4). Thus, the predator population is composed of both specialist and flexible moving individuals, and individuals further distinguish themselves as specialist/flexible hunters with experience (Figure 2).

@@

At all experience levels, predators that were on average faster (mean) were also travelling faster in all their matches (IIV) (Figure 3-S1). Thus, faster predators specialized at travelling fast, while slower predators were more flexible in their movement through time. In general, predators tended to be faster (mean) when they encountered prey that were on average faster (mean) (Figure 3-S1). At the three experience levels, there were only weak correlations between predator IIV and mean prey speed (Figure 3-S1). Similarly, predators specialized in a faster foraging mode (IIV) when encounters with prey were more predictable (IIV), but the correlations were weak (Figure 3-S1). In the four cases described above, the correlations became stronger as predators gained experience (Figure 3-S1).

The foraging strategies of predators, that is, the correlations between the mean and IIV of speed, remained structurally stable among the levels of experience. For instance, novice predators that tended to be faster also kept this strategy as they gained experience (Figure 3 inside the black squares). Overall, we detected the same patterns for within-experience state correlations. The correlations between trait values were also stronger among novice-intermediate and intermediate-advanced states than between novice-advanced states, indicating that behavioural changes in predators occurred as they gained experience.

```{r fig3, out.width = "100%", fig.cap = "**Figure 3.** Full correlation matrix of the MDHGLM relating the mean and IIV speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Starting from the upper part of the plot are displayed the five parameters for novice predators (gray), the five for intermediate predators (yellow), and the five for advanced predators (blue). Larger dark blue circles indicate stronger positive correlations, while larger dark red circles indicate stronger negative correlations. The black squares indicate correlations between different experience levels (i.e. character states)."}

@@

We did not detect strong correlations between the predators' average speed and hunting success (Figure 3-S1), nor between the IIV in speed and hunting success (Figure 3-S1), suggesting that both slower/flexible and faster/specialist predators were both equally successful. However, when predators were novice, there was a strong negative correlation between the average speed of the prey that they encountered and their hunting success (Figure 3-S1). Hence, novice individuals that encountered faster prey were less successful. This relationship was decoupled at the later experience levels, such that encounters with faster prey at intermediate and advanced stages were not as strongly correlated with hunting success anymore (Figure 3-S1).

@@

## Experience strenghens predator-prey interactions

>>>>

For each level of experience (i.e. novice, intermediate, advanced), the model estimated the mean speed of every individual (which we call the mean part of the model), and, employing a heterogeneous structure of the residuals, the intra-individual standard deviation in speed of every individual (which we call the dispersion part of the model). Thus, in the mean part of the model, we could estimate among individual differences in the foraging tactic used by predators (i.e. fast or slow), while in the dispersion part of the model, we could estimate among individual differences in foraging specialization/flexibility (i.e. intra-individual variability). We followed the same structure for the preys' speed to estimate among individual differences in the mean and standard deviation in speed of the prey encountered by predators. For hunting success, we only modeled the mean part of the equation to estimate among individual differences in mean hunting success. For the predator and the prey's speed, we controlled for the average rank of the prey that they encountered on both the mean and dispersion part of the equation. The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank). We used the mean rank of the four prey and applied a square-root transformation. For the predator's hunting success, we controlled for the square-root of the duration of the match. The model has the following form, with each equation repeated three times (i.e. one for each level of experience) :

@@

At the individual level, we detected differences in the average movement speed among predators ($SD\ predator\ ID$, Table 2). These differences remained stable (although they decreased minimally) at the intermediate stage (Table 2 and Figure S4), and increased at the advanced stage to a level higher than when predators were novice (Table 2 and Figure S4). For the dispersion part of the model (i.e. sigma predator speed), predators displayed considerable differences in their degree of hunting specialization, which also increased with experience ($SD\ predator\ ID$, Table 2 and Figure S4). Thus, the predator population is composed of both specialist and flexible moving individuals, and they further distinguish themselves as specialist/flexible hunters as they gain experience (Figure 2).

@@

At all experience levels, predators that were on average faster (mean) had also a lower intra-individual variability (IIV) across their matches (Figure 3 and Table S1). Thus, faster predators specialized at travelling fast, while slower predators were more flexible in their movement through time. In general, predators that encountered prey that were on average faster (mean) tended to be faster (mean) (Figure 3 and Table S1). At the three experience levels, there were only weak correlations between predator IIV and mean prey speed (Figure 3 and Table S1). However, predators that experienced more predictable encounters with their prey (IIV) specialized in a faster hunting tactic (IIV) (Figure 3 and Table S1). For all the results described above, the correlations became stronger as predators gained experience (Figure 3 and Table S1).

The foraging strategies of predators, that is, the correlations between the mean and IIV of speed, remained structurally stable among the levels of experience (Figure 3 inside the black circles and Table S1). For instance, novice predators that tended to be faster also kept this strategy as they gained experience (Figure 3 inside the black squares and Table S1). Overall, we detected the same patterns than those from the within-experience state correlations described above. The correlations between trait values were also stronger among novice-intermediate and intermediate-advanced states than between novice-advanced states, indicating that the behavioural tactics of predators were more similar among experience levels closer in time (Figure 3 inside the black circles and Table S1).

```{r fig3, out.width = "100%", fig.cap = "**Figure 3.** Full correlation matrix of the MDHGLM relating the mean and intra individual variability (IIV) in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Starting from the upper part of the plot are displayed the five parameters for novice predators (gray), the five for intermediate predators (yellow), and the five for advanced predators (blue). Larger dark blue circles indicate stronger positive correlations, while larger dark red circles indicate stronger negative correlations. The black squares indicate correlations between different experience levels (i.e. character states)."}

@@

We did not detect strong correlations between the predators' average speed and their hunting success (Figure 3 and Table S1), nor between the predators' IIV in speed and their hunting success (Figure 3 and Table S1), suggesting that both slower/flexible and faster/specialist predators were both equally successful. However, when predators were novice, there was a strong negative correlation between the average speed of the prey that they encountered and their hunting success (Figure 3 and Table S1). Hence, novice individuals that encountered faster prey were less successful. This relationship was decoupled at the later experience levels, such that encounters with faster prey at intermediate and advanced stages were not as strongly correlated with hunting success anymore (Figure 3 and Table S1).

@@

## Experience strenghtens predator-prey interactions
